---
title: "NABat_Annual_Report_Appendices"
author: "Created by Joanna Burgar"
date: "Report generated on `r format(Sys.time(), '%d %B, %Y')`"
output: word_document

---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE)

#Load Packages
list.of.packages <- c("data.table", "leaflet", "tidyverse", "lunar", "zoo", "colortools", "lubridate", "camtrapR", "circular", "RColorBrewer", "Cairo", "viridis", "knitr", "sf","osmdata", "ggspatial", "ggmap","gridExtra", "grid")
# Check you have them and load them
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(list.of.packages, require, character.only = TRUE)

load("NABat_Annual_Report.RDS")
source("Appendix_maps.R", local = knitr::knit_global())

```

## NABat 2020 Report Appendix

``` {r map with inset, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=9}
  grid.newpage()
  vpb_ <- viewport(width = 1, height = 1, x = 0.4, y = 0.5)  # the larger map
  vpa_ <- viewport(width = 0.6, height = 0.5, x = 0.85, y = 0.8)  # the inset in upper right
  print(GRTS.plot(GRTS.Cell.ID=params$GRTS.Cell.ID, h.just=0), 
      vp = vpb_)
  print(ggplot() + 
        geom_sf(data = Alberta) +
        geom_sf(data = NR, mapping = aes(fill=NR), lwd=0) + 
        scale_fill_manual(name = "Natural Regions",
                          values=c("#669933","cadetblue3","#CCFF99","#FFCC66","chocolate1","#CC3333"))+
        geom_sf(data = NABat_grid, col="azure2", lwd=0.8) +
        geom_sf(data = NABat_grid %>% dplyr::filter(GRTS_ID==params$GRTS.Cell.ID), col="black", lwd=1) +
        #annotation_scale(location = "bl",bar_cols = c("grey", "white")) +
        coord_sf() +
        theme(axis.text.x = element_text(size=4), axis.text.y =element_text(size=5))+
        theme(legend.position = "none"),
      vp = vpa_)
```

##### Figure 1. GRTS Cell `r params$GRTS.Cell.ID` (black square) with stations survyed in `r unique(dat_count$Deployment.ID)` and inset map showing grid location compared to other grid cells (gray) surveyed in `r unique(dat_count$Deployment.ID)` throughout Alberta. 


##### Table 1. NABat station meta data for GRTS Cell `r params$GRTS.Cell.ID`.  

``` {r Table 1, echo=FALSE, warning=F}
###- Appendix Table 1 (formatting/structuring in Appendix_read)
  opts <- options(knitr.kable.NA = "")
  knitr::kable(t(Appendix.Table1 %>% dplyr::filter(GRTS.Cell.ID==params$GRTS.Cell.ID) %>% select(-GRTS.Cell.ID)),
             #caption = paste("Table 1 - NABat Site Metadata for GRTS Cell",params$GRTS.Cell.ID),
             digits=0)
  
```


##### Table 2. Minimum and maximum nightly survey weather conditions - Temperature (°C), Humidity (%), and Wind Speed (km/hr) - for GRTS Cell `r params$GRTS.Cell.ID`.  

``` {r Table 2, echo=F, warning=F}
  
###- Appendix Table 2
  knitr::kable(Appendix.Table2 %>% dplyr::filter(grepl(params$GRTS.Cell.ID, Location.Name)) %>% select(-Location.Name),
             caption = paste("Table 2 - NABat Survey Weather Conditions for GRTS Cell",params$GRTS.Cell.ID),
             digits=1)
```



``` {r Figure 2, echo=FALSE, warning=FALSE, results=FALSE}
###- Appendix Figure 2
  # only graph species / species groups with >100 calls
sp.to.use <- call_count.Sp %>% group_by(GRTS.Cell.ID, Classification) %>% summarise(Sum = sum(Count)) %>% filter(Sum>100)
  
  
  # group data for overall mean nightly bat calls by year with volancy
app.Calls <- call_count.Sp%>% group_by(GRTS.Cell.ID, Classification, Deployment.ID) %>% summarise(Mean = mean(Count), SE = se(Count))
  
  
  app.calls.Sp <- app.Calls %>% dplyr::filter(GRTS.Cell.ID==params$GRTS.Cell.ID) %>% 
    filter(Classification %in% sp.to.use$Classification) %>%
    ggplot(aes(x = Classification, y = Mean, fill="GRTS.Cell.ID"))+
    geom_point(colour="white", shape=21, size=4)+
    scale_fill_manual(values=unique(col.catVol[1])) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    ylab(expression(paste("Mean ± 1 SE Nightly Bat Calls"))) +
    geom_linerange(aes(Classification, ymin = Mean-SE, ymax = Mean+SE)) +
    theme_classic()+
    theme(axis.text.y = element_text(size=12), axis.title.x = element_blank())+
    theme(axis.text.x = element_text(colour = "black", size = 10)) +
    theme(legend.position = "none", legend.title = element_blank())+
    facet_wrap(~Deployment.ID)
  
  app.calls.Sp
  
```  
  
##### Figure 2. Mean (± 1 Standard Error) nightly bat calls for GRTS Cell `r params$GRTS.Cell.ID` in `r unique(dat_count$Deployment.ID)`. 
  
``` {r Figure 3, echo=FALSE, warning=FALSE}
###- Appendix Figure 3
  
app.hist.calls <- ggplot(data = call_count.Sp %>% dplyr::filter(GRTS.Cell.ID==params$GRTS.Cell.ID), aes(x = Classification, y = Count, fill= Location.Name)) + 
    geom_bar(stat = "identity") +
    scale_fill_brewer(palette = "Paired")+
    theme_classic() + xlab("Species / Species Group") + ylab("Total Bat Calls") +
    theme(legend.position="none") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, colour = "black", size = 10)) +
    theme(axis.title.y = element_text(size = 14)) +   theme(axis.title.x = element_blank())+
    facet_wrap(~Deployment.ID)
  
app.hist.calls
  
```

##### Figure 3. Total bat calls for GRTS Cell `r params$GRTS.Cell.ID` in `r unique(dat_count$Deployment.ID)`. 


##### Table 3. Nightly call counts for GRTS Cell `r params$GRTS.Cell.ID`, as identified by Alberta eBat to species / species groups, unknown, or as noise (i.e., not a bat).

``` {r Table 3, echo=FALSE, warning=FALSE}
###- Appendix Table 3 - Nightly call counts
  Appendix.Table3 <- dat_count %>% group_by(Location.Name, SurveyNight, Classification) %>%
    summarise(Call.Count = sum(Count))
  
  Appendix.Table3 <- Appendix.Table3 %>% pivot_wider(names_from = Classification, values_from = Call.Count)
  Appendix.Table3 <- Appendix.Table3[,c("Location.Name","SurveyNight","EPFU", "EPFU-LANO", "LANO", "LACI", "LABO", "LABO-MYLU", "MYLU","MYCA", "MYCI", "MYEV","MYEV-MYSE", "MYSE", "MYVO", "Myotis 40k", "unknown","noise")]
  colnames(Appendix.Table3)[1:2] <- c("Station", "Date")
  
  
  knitr::kable(Appendix.Table3 %>% dplyr::filter(grepl(params$GRTS.Cell.ID, Station)),
             #caption = paste("Table 3 - NABat Nightly Call Counts for GRTS Cell",i),
             digits=1)
options(opts)


```
